### 2014-10-15 PrivacyProtection-2014-09-09.tex
Updated IT Systems description

In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State authority.

@@

The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA.
Although the act seems to fully cover the directive.
Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e. consent must be given freely and has to be explicit.

The Act's definition of personal data covers any data that can be used to identify a living individual.
Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.

The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.

The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
The person who has their data processed has the right to:
\begin{itemize}
\item View the data an organization holds on them.
A `subject access request' can be obtained for a nominal fee.
As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.
\item Request that incorrect information be corrected.
If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.
\item Require that data is not used in any way that may potentially cause damage or distress.
\item Require that their data is not used for direct marketing. \cite{3}
\end{itemize}

@@

@@

A citizen (a) carries a mobile device running the \emph{sensor collector} (d) application.
The sensor collector application collects sensor data (accelerometer, GPS, GSM, ...) and sends it to the \emph{data center} (c).
The sensor collector can also be able to perform certain data mining operations.
Examples of such data mining operations include human activity recognition,the detection of service lines, detection of characters and faces from images.

The \emph{data center} stores and processes sensor data collected with the sensor collector application.
Again the amount

Vice versa, it sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
The \textit{System Provider (b)} provides and operates technical infrastructure like the \textit{Data Center} and the \textit{Report Tool (f)}.
The \textit{Report Tool} queries the \textit{Data Center} for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored \textit{Citizens}. 
\textit{Local Authorities (c)} use the \textit{Report Tool} to get information in order to understand citizen movement and improve public services.

In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining} (Human Activity Recognition: detect running, walking, standing, ...). 
Usually, \textit{Data Mining} is heavily conducted within the \textit{Data Center}, but simpler mining processes can also be conducted on mobile devices.

>>>>

The Live+Gov systems can violate this fundamental human right if it is used to gather information about the private life without the knowledge and consent of the citizen.

% In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State authority.

@@

%TODO: Reformulate Wikipedia Text

%% The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA.
%% Although the act seems to fully cover the directive.
%% Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e. consent must be given freely and has to be explicit.

%% The Act's definition of personal data covers any data that can be used to identify a living individual.
%% Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.

%% The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.

%% The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
%% The person who has their data processed has the right to:
%% \begin{itemize}
%% \item View the data an organization holds on them.
%% A `subject access request' can be obtained for a nominal fee.
%% As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.
%% \item Request that incorrect information be corrected.
%% If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.
%% \item Require that data is not used in any way that may potentially cause damage or distress.
%% \item Require that their data is not used for direct marketing. \cite{3}
%% \end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%

@@

A \emph{citizen} (a) carries a mobile device running the \emph{sensor collector} application (d).
The sensor collector application is able to collect various kinds of sensor data (accelerometer, GPS, GSM, ...) and can sent the raw data back to the \emph{data center} (c).
The sensor collector can also be able to perform certain data mining operations.

Examples of such data mining operations include human activity recognition,the detection of service lines, detection of characters and faces from images.

In the Live+Gov Project mobile devices are used in particular for the following activities:
\begin{enumerate}
\item collection of GPS samples,
\item mining of Human activities (HAR) based on accelerometer samples,
\item collection of reports consisting of an image, free text and selected categories.
\end{enumerate}

The \emph{data center} stores and processes sensor data collected with the sensor collector application. It can also take into account data obtained from third parties, like the current positions of trains.
The data center can sent mining end products (traffic jam reports, bus schedule) back to the mobile device of the citizen.

In the Live+Gov Project data centers, in particular, perform the following activities:
\begin{enumerate}
\item storage of login credentials, name, and email address for each citizen,
\item storage of collected GPS samples (user id, timestamp, GPS location),
\item storage of HAR results (user id, timestamp, HAR result),
\item detection of service lines based received GPS samples (SLD),
\item storage of SLD results (userid, timestamp, SLD result),
\item sent back SLD results to mobile device,
\item storage of received reports (user id, timestamp, report),
\item detection of inherent patterns of the received reports.
\end{enumerate}

The \emph{System Provider} (b) provides and operates technical infrastructure like the Data Center and the \emph{Reporting Tool} (f).
The Reporting Tool queries the Data Center for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored citizens.

\textit{Local Authorities} (c) use the reporting tool to get information in order to understand citizen movement and improve public services.
In such systems, the most valuable information for local authorities is not the raw sensor data, but aggregated views on the mining end products.

In the Live+Gov Project the reporting tools allows, in particular the following queries:
\begin{enumerate}
\item show aggregate information about which routes citizens take starting from a given location,
\item show the average waiting time of a citizen for each bus stop,
\item show routes where citizens where running to catch a bus,
\item show locations of all reports in a given time window,
\item visualize detected patterns in the report data,
\item view individual reports.
\end{enumerate}
